### 2014-04-23 manuscript.tex
basic structure based off elsarticle template

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

%\journal{Journal of \LaTeX\ Templates}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
\bibliographystyle{model2-names}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Phylogeny, divergence times, and chromosome evolution in Onagraceae}

\author[berk]{William A. Freyman\corref{cor1}}
\ead{freyman@berkeley.edu}
\cortext[cor1]{Corresponding author}

\address[berk]{1600 John F Kennedy Boulevard, Philadelphia}

\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

%\begin{keyword}
%\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
%\MSC[2010] 00-01\sep  99-00
%\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}

blah blah

\section{Methods}

\paragraph{Supermatrix} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.

\paragraph{Phylogenetic analysis} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript. Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal. It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.

\paragraph{Chromosome evolution} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class. In addition, it features commands and options to format the
\begin{itemize}
\item document style
\item baselineskip
\item front matter
\item keywords and MSC codes
\item theorems, definitions and proofs
\item lables of enumerations
\item citation style and labeling.
\end{itemize}

\section{Results}

The author names and affiliations could be formatted in two ways:
\begin{enumerate}[(1)]
\item Group the authors per affiliation.
\item Use footnotes to indicate the affiliations.
\end{enumerate}
See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.

\section{Conclusion}

There are various bibliography styles available. You can select the style of your choice in the preamble of this document. These styles are Elsevier styles based on standard styles like Harvard and Vancouver. Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.

Here are two sample references: \citep{Feynman1963118,Dirac1953888}.

\section*{References}

\bibliography{manuscriptbib}

\end{document}

### 2014-04-24 manuscript.tex
cleaning up some

@@
 
-%\journal{Journal of \LaTeX\ Templates}
+\graphicspath{ {./figures/} }
+
+\journal{IB200}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%
-%% To change the style, put a % in front of the second line of the current style and
-%% remove the % from the second line of the style you would like to use.
-%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Numbered
-%\bibliographystyle{model1-num-names}
-
-%% Numbered without titles
-%\bibliographystyle{model1a-num-names}
 
@@
 
-%% Vancouver numbered
-%\usepackage{numcompress}\bibliographystyle{model3-num-names}
-
-%% Vancouver name/year
-%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
-
-%% APA style
-%\bibliographystyle{model5-names}\biboptions{authoryear}
-
-%% AMA style
-%\usepackage{numcompress}\bibliographystyle{model6-num-names}
-
 %% `Elsevier LaTeX' style
 %\bibliographystyle{elsarticle-num}
+
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\title{Phylogeny, divergence times, and chromosome evolution in Onagraceae}
+\title{Fossil-calibrated phylogeny and chromosome evolution of Onagraceae}
 
@@
 
-\address[berk]{1600 John F Kennedy Boulevard, Philadelphia}
+\address[berk]{Jepson Herbarium and Department of Integrative Biology, University of California, Berkeley}
 
@@
 
-\paragraph{Supermatrix} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.
+\paragraph{Supermatrix}
+
+1: Downloads GenBank database of the specified GB division (PLN, MAM, etc)
+2: Perform exhaustive all-by-all BLAST comparisons of each ingroup and outgroup sequence.
+3: Alternatively (much faster) uses a FASTA file of guide sequences to define each cluster.
+   Each ingroup/outgroup sequence is BLASTed against the guide sequences.
+   4: Build clusters of sequences:
+           - use single-linkage hierarchical clustering algorithm
+	           - distance threshold default BLASTn e-value 1.0e-10
+		           - uses sequence length percent similarity cutoff default 0.5
+			           - discards clusters that are not phylogenetically informative (< 4 taxa)
+				   5: Aligns each cluster of sequences using MUSCLE.
+				   6: Concatenates the clusters creating a supermatrix.
 
@@
 
+\begin{figure*}[t]
+    \centering
+    % have to trim bottom off image
+    \includegraphics[width=1.0\textwidth, trim=0 10 0 0, clip=true]{colored_posterior}
+    \caption{Phylogeny blah blah blah}
+    \label{posteriors}
+\end{figure*}
+
+\begin{figure*}[t]
+    \centering
+    % have to trim bottom off image
+    \includegraphics[width=1.0\textwidth, trim=0 10 0 0, clip=true]{time_colored_genera}
+    \caption{Phylogeny blah blah blah}
+    \label{genera}
+\end{figure*}
+
 \section{Conclusion}
 
-There are various bibliography styles available.
+There are various bibliography styles availablei, please see Figure \ref{genera}.
 You can select the style of your choice in the preamble of this document.

### 2014-04-26 manuscript.tex
new figures and stuff

@@
 
-\title{Fossil-calibrated phylogeny and chromosome evolution of Onagraceae}
+\title{Fossil-calibrated phylogeny of Onagraceae: patterns of floral and chromosome evolution}
 
@@
 
+Here is a table of the fossil calibrations used....
+
 \paragraph{Chromosome evolution} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class.
@@
 
+Here is a table with the mean divergence time estimates including \%95 confidence intervals....
+
 The author names and affiliations could be formatted in two ways:
@@
 
-\begin{figure*}[t]
-    \centering
+\begin{figure*}[p]
+    \vspace*{-2cm}
+    \makebox[\linewidth]{
         % have to trim bottom off image
-    \includegraphics[width=1.0\textwidth, trim=0 10 0 0, clip=true]{colored_posterior}
+        \includegraphics[width=1.6\linewidth, trim=0 10 0 0, clip=true]{colored_posterior}
+    }
     \caption{Phylogeny blah blah blah}
@@
 
-\begin{figure*}[t]
-    \centering
-    % have to trim bottom off image
-    \includegraphics[width=1.0\textwidth, trim=0 10 0 0, clip=true]{time_colored_genera}
-    \caption{Phylogeny blah blah blah}
+\begin{figure*}[p]
+    \vspace*{-2cm}
+    \makebox[\linewidth]{
+        \includegraphics[width=1.6\linewidth, trim=0 10 0 0, clip=true]{time_colored_genera}
+    }
+    \caption{Sequence diagram}
     \label{genera}

### 2014-04-27 manuscript.tex
much of methods section, reformatting throughout

@@
 
-\usepackage{lineno,hyperref}
+\usepackage{lineno}
+\usepackage[hidelinks]{hyperref}
+\usepackage{mathtools}
+\usepackage{adjustbox}
+
 \modulolinenumbers[5]
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%
-%% Elsevier bibliography styles
-%%%%%%%%%%%%%%%%%%%%%%%
-
+%% bibliography styles
 %% Harvard
 \bibliographystyle{model2-names}\biboptions{authoryear}
-
 %% `Elsevier LaTeX' style
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%
+%% front matter
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{abstract}
-This template helps you to create a properly formatted \LaTeX\ manuscript.
+blah blah
 \end{abstract}
 
-%\begin{keyword}
-%\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
-%\MSC[2010] 00-01\sep  99-00
-%\end{keyword}
+\end{frontmatter}
 
-\end{frontmatter}
+%%%%%%%%%%%%%%%%%%%%%%%
+%% intro
+%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\paragraph{Supermatrix}
+\paragraph{Supermatrix assembly}
 
-1: Downloads GenBank database of the specified GB division (PLN, MAM, etc)
-2: Perform exhaustive all-by-all BLAST comparisons of each ingroup and outgroup sequence.
-3: Alternatively (much faster) uses a FASTA file of guide sequences to define each cluster.
-   Each ingroup/outgroup sequence is BLASTed against the guide sequences.
-   4: Build clusters of sequences:
-           - use single-linkage hierarchical clustering algorithm
-	           - distance threshold default BLASTn e-value 1.0e-10
-		           - uses sequence length percent similarity cutoff default 0.5
-			           - discards clusters that are not phylogenetically informative (< 4 taxa)
-				   5: Aligns each cluster of sequences using MUSCLE.
-				   6: Concatenates the clusters creating a supermatrix.
+I downloaded all available DNA sequences from GenBank release 200 PLN division and
+performed an exhaustive all-by-all BLASTn \citep{blast} comparison for sequences in Onagraceae
+and Lythraceae.
+Using a BLASTn e-value of $1.0e^{-10}$ threshold and a sequence length
+percent similarity cutoff of 0.5,
+I constructed clusters of putative homologs using a single-linkage hierarchical clustering algorithm.
+Subspecies names were removed from all sequences, and all but one sequence of each species was pruned from each cluster.
+Clusters that were not phylogenetically informative ($< 4$ taxa) were discarded,
+and each cluster was aligned using MUSCLE \citep{edgar2004muscle}. 
+The alignments were concatenated by species, and any species that was not present in at least
+two clusters was removed from the supermatrix.
+Code written to assemble the supermatrix is availabe as the Python module 
+Supermatrix Constructor (SUMAC) \citep[\url{https://github.com/wf8/sumac}]{sumac}.
 
-\paragraph{Phylogenetic analysis} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript.
-Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal.
-It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.
+\paragraph{Phylogenetic analyses} 
+Maximum likelihood analyses were performed with RAxML-HPC \citep{raxml} on the CIPRES Scientific Gateway \citep{cipres} 
+using the rapid bootstrap heuristic and the GTRCAT nucleotide substitution model.
+I used the ML tree to select 15 species phylogenetically widely distributed in Lythraceae to act as outgroup; 
+all other members of Lythraceae were subsequently removed from the supermatrix before estimating divergence times.
+Bayesian estimates of divergence times were inferred using BEAST \citep{beast, beast2} on CIPRES and calibrated with five fossils 
+identified with morphological synapomorphies (Table \ref{fossils}).
+The \textit{Ludwigia} fossil pollen was dated broadly to the Paleocene \citep{grimsson}, so I set the prior to a normal distribution with a wide 
+standard deviation to cover the entire time period.
+For all other calibration points I used a lognormal prior distribution with the offset (the minimum age of the node) corresponding to the fossil age.
+The BEAST analysis utilized the GTR+$\Gamma$ nucleotide substitution model with a relaxed molecular clock (uncorrelated lognormal model)
+and a Yule process tree prior.
+The Markov Chain Monte Carlo (MCMC) was run for 100 million generations, sampling every 10 thousand generations.
+Tracer v1.6 \citep{tracer} was used to assess the MCMC output for parameter convergence and ensure that the effective sample size for all parameters was above 200.
+The first 1000 trees were discarded as burn-in, and the remaining 9000 trees were summarized as a maximum clade credibility tree with mean divergence times.
 
-Here is a table of the fossil calibrations used....
+\paragraph{Character evolution}
+blah blah
 
-\paragraph{Chromosome evolution} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class.
-In addition, it features commands and options to format the
-\begin{itemize}
-\item document style
-\item baselineskip
-\item front matter
-\item keywords and MSC codes
-\item theorems, definitions and proofs
-\item lables of enumerations
-\item citation style and labeling.
-\end{itemize}
+\begin{table*}
+   \begin{adjustbox}{max width=\textwidth}
+      \begin{tabular}{lllllll}
+         \hline
+         Group & Age (Mya) & Prior Distribution & Mean & SD & Offset & Reference \\ \hline
+         \textit{Circaea} (Onagraceae) & 12 & lognormal & 0.0 & 2.0 & 12 & \citep{grimsson} \\
+         \textit{Epilobium} (Onagraceae) & 12 & lognormal & 0.0 & 2.0 & 12 & \citep{grimsson} \\
+         S. Pacific \textit{Fuschia} (Onagraceae) & 23 & lognormal & 0.0 & 1.0 & 23 & \citep{lee2013fossil} \\
+         \textit{Ludwigia} (Onagraceae) & Paleocene & normal & 60.0 & 3.0 & - & \citep{zhi} \\
+         Lythraceae & 82 & lognormal & 0.0 & 2.0 & 82 & \citep{graham} \\
+         \hline
+      \end{tabular}
+   \end{adjustbox}
+   \caption{Fossil priors used in BEAST analyses.}
+   \label{fossils}
+\end{table*}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+%% results
+%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+\paragraph{Supermatrix assembly}
+blah blah
+
+\paragraph{Divergence time estimates}
 Here is a table with the mean divergence time estimates including \%95 confidence intervals....
 
-The author names and affiliations could be formatted in two ways:
-\begin{enumerate}[(1)]
-\item Group the authors per affiliation.
-\item Use footnotes to indicate the affiliations.
-\end{enumerate}
-See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.
+\paragraph{Character evolution}
+blah blah
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%
+%% conclusion
+%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Conclusion}
 
-There are various bibliography styles availablei, please see Figure \ref{genera}.
-You can select the style of your choice in the preamble of this document.
-These styles are Elsevier styles based on standard styles like Harvard and Vancouver.
-Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.
+blah blah
 
-Here are two sample references: \citep{Feynman1963118,Dirac1953888}.
+%%%%%%%%%%%%%%%%%%%%%%%
+%% references
+%%%%%%%%%%%%%%%%%%%%%%%
 

### 2014-04-28 manuscript.tex
character state recon. started

@@
 
-\paragraph{Character evolution}
-blah blah
-
 \begin{table*}
@@
 \end{table*}
+
+\paragraph{Character state reconstruction}
+I scored six characters, including
+chromosome number, floral merosity, petal color, and self-compatibility/incompatibility. 
+Character data was assembled from the comprehensive \citet{wagner2007revised} Onagraceae monograph.
+Ancestral character state reconstructions of petal number and petal color were performed using Mesquite v2.75 \citep{mesquite}.
+Characters were treated as unordered categorical data, and optimized using maximum likelihood
+under the Markov k-state 1 parameter (Mk1) model \citep{lewis2001likelihood}.
 

### 2014-04-28 manuscript.tex
added d. time table, and many other changes

--- 2014-04-27 manuscript.tex
@@
 two clusters was removed from the supermatrix.
-Code written to assemble the supermatrix is availabe as the Python module 
-Supermatrix Constructor (SUMAC) \citep[\url{https://github.com/wf8/sumac}]{sumac}.
+Code written to assemble the supermatrix is available as the Python module 
+SUMAC \citep[\url{https://github.com/wf8/sumac}]{sumac}, and can be used to assemble
+supermatrices for other taxonomic groups recognized in GenBank.
 
@@
 using the rapid bootstrap heuristic and the GTRCAT nucleotide substitution model.
-I used the ML tree to select 15 species phylogenetically widely distributed in Lythraceae to act as outgroup; 
-all other members of Lythraceae were subsequently removed from the supermatrix before estimating divergence times.
-Bayesian estimates of divergence times were inferred using BEAST \citep{beast, beast2} on CIPRES and calibrated with five fossils 
+I used the ML tree to select 15 taxa phylogenetically widely distributed in Lythraceae to act as outgroup for the divergence time analysis; 
+all other members of Lythraceae were subsequently removed from the supermatrix.
+Bayesian estimates of divergence times were inferred using BEAST v1.8 \citep{beast, beast2} on CIPRES and calibrated with five fossils 
 identified with morphological synapomorphies (Table \ref{fossils}).
@@
 Tracer v1.6 \citep{tracer} was used to assess the MCMC output for parameter convergence and ensure that the effective sample size for all parameters was above 200.
-The first 1000 trees were discarded as burn-in, and the remaining 9000 trees were summarized as a maximum clade credibility tree with mean divergence times.
-
-\paragraph{Character evolution}
-blah blah
+The first 1000 trees were discarded as burn-in, and the remaining 9000 trees were summarized as a maximum clade credibility (MCC) tree with mean divergence times.
 
@@
          \hline
-         Group & Age (Mya) & Prior Distribution & Mean & SD & Offset & Reference \\ \hline
+         Group & Age (Mya) & Prior Distribution & Mean & SD & Offset & Reference \\ 
+	 \hline
          \textit{Circaea} (Onagraceae) & 12 & lognormal & 0.0 & 2.0 & 12 & \citep{grimsson} \\
@@
    \end{adjustbox}
-   \caption{Fossil priors used in BEAST analyses.}
+   \caption{Fossils used as priors in the Bayesian divergence time analysis.}
    \label{fossils}
 \end{table*}
+
+\paragraph{Character state reconstruction}
+I scored six characters, including
+chromosome number, floral merosity, petal color, and self-compatibility/incompatibility. 
+Character data was assembled from the comprehensive \citet{wagner2007revised} Onagraceae monograph.
+Ancestral character state reconstructions of petal number and petal color were performed using Mesquite v2.75 \citep{mesquite}
+over the Bayesian MCC tree.
+Characters were treated as unordered categorical data, and optimized using maximum likelihood
+with the Markov k-state 1 parameter (Mk1) model \citep{lewis2001likelihood}.
 
@@
 Here is a table with the mean divergence time estimates including \%95 confidence intervals....
+\begin{table*}
+   \begin{adjustbox}{max width=\textwidth}
+      \begin{tabular}{lllllll}
+         \hline
+         Clade & Mean Age (Mya) & \%95 HPD Min & \%95 HPD Max \\ 
+	 \hline
+         Onagraceae / Lythraceae & 109 & 88 & 131 \\
+         \textit{Ludwigia} & 97 & 76 & 118 \\
+	 \textit{Hauya} & 49 & 35 & 64 \\
+	 \textit{Circaea} / \textit{Fuchshia} & 37 & 28 & 47 \\
+	 \textit{Lopezia} & 71 & 55 & 68 \\
+	 \textit{Gongylocarpus} & 60 & 45 & 77 \\
+	 \textit{Epilobium} & 49 & 38 & 60 \\
+	 \textit{Chamerion} & 47 & 36 & 57 \\
+	 \textit{Xylonagra} & 43 & 33 & 52 \\
+	 \textit{Clarkia} & 40 & 32 & 48 \\
+         \textit{Terapteron} & 19 & 10 & 29 \\
+	 \textit{Camissoniopsis} / \textit{Neoholmgrenia} & 14 & 5 & 23 \\
+	 \textit{Eremothera} / \textit{Camissonia} & 24 & 16 & 33 \\
+	 \textit{Taraxia} & 30 & 22 & 38 \\
+	 \textit{Chylismiella} / \textit{Gayophytum} & 20 & 10 & 30 \\
+	 \textit{Eulobus} & 26 & 19 & 34 \\
+	 \textit{Chylismia} / \textit{Oenothera} & 25 & 18 & 31 \\
+         \hline
+      \end{tabular}
+   \end{adjustbox}
+   \caption{Bayesian divergence time estimates of major clades.}
+   \label{times}
+\end{table*}
 
@@
     }
-    \caption{Phylogeny blah blah blah}
+    \caption{
+       Bayesian maximum clade credibility phylogeny of 280 Onagraceae taxa and 15 Lythraceae taxa. 
+       Estimated posterior probabilities close to 1.0 are shown in green. All genera described in
+       \citet{wagner2007revised} are monophyletic with posterior probabilities of $> 0.95$
+       except for sister genera \textit{Neoholmgrenia} and \textit{Camissoniopsis} (posterior = 0.31).
+    }
     \label{posteriors}
@@
     }
-    \caption{Sequence diagram}
+    \caption{
+       Bayesian chronogram of 280 Onagraceae taxa and 15 Lythraceae taxa.
+       Approximate positions of fossil calibration points are shown as black circles.
+       All genera described in \citet{wagner2007revised} are colored, and their
+       divergence time estimates and \%95 HPD intervals can be seen in Table \ref{times}.
+}
     \label{genera}

### 2014-04-28 manuscript.tex
added lots to results

@@
 and Lythraceae.
-Using a BLASTn e-value of $1.0e^{-10}$ threshold and a sequence length
+Using a BLASTn e-value of $1.0 \times 10^{-10}$ threshold and a sequence length
 percent similarity cutoff of 0.5,
@@
 \paragraph{Phylogenetic analyses} 
-Maximum likelihood analyses were performed with RAxML-HPC \citep{raxml} on the CIPRES Scientific Gateway \citep{cipres} 
+Maximum likelihood (ML) analyses were performed with RAxML-HPC \citep{raxml} on the CIPRES Scientific Gateway \citep{cipres} 
 using the rapid bootstrap heuristic and the GTRCAT nucleotide substitution model.
@@
 \paragraph{Supermatrix assembly}
-blah blah
+SUMAC evaluated 5571 Onagraceae and 2832 Lythraceae nucleotide sequences to construct the supermatrix. 
+The completed supermatrix consisted of 11 clusters of homologous sequences (Table \ref{clusters}).
+As used in the maximum likelihood analyses (before pruning the number of outgroup taxa), 
+the supermatrix contained 521 taxa, was 31862 nucleotides long, and contained 93.0\% missing data.
 
-\paragraph{Divergence time estimates}
-Here is a table with the mean divergence time estimates including \%95 confidence intervals....
 \begin{table*}
@@
          \hline
-         Clade & Mean Age (Mya) & \%95 HPD Min & \%95 HPD Max \\ 
+         DNA Region & \# of Taxa & Aligned Length & Missing data (\%) & Taxon Coverage Density \\ 
+	 \hline
+         ITS & 453 & 1746 & 13.2 & 0.87 \\
+         trnL & 234 & 1429 & 55.2 & 0.45 \\
+         rpl16 & 91 & 1414 & 82.6 & 0.17 \\
+         rbcL & 77 & 1474 & 85.2 & 0.15 \\
+         rps16 & 74 & 1016 & 85.8 & 0.14 \\
+         rbcL & 64 & 1310 & 87.7 & 0.12 \\
+         PgiC2 & 47 & 4028 & 91.0 & 0.09 \\
+         matK & 37 & 921 & 92.9 & 0.07 \\
+         ndhF & 37 & 2063 & 92.9 & 0.07 \\
+         pgiC & 26 & 14709 & 95.0 & 0.05 \\
+         R5 & 18 & 3129 & 96.6 & 0.03 \\
+         \hline
+      \end{tabular}
+   \end{adjustbox}
+   \caption{Clusters of homologous sequences used to assemble the supermatrix.}
+   \label{clusters}
+\end{table*}
+
+\paragraph{Phylogeny and divergence time estimates}
+The topologies of the ML and Bayesian phylogenies were identical for all major clades within Onagraceae, 
+so only the Bayesian MCC tree (Figures \ref{posteriors} and \ref{genera}) is shown here.
+All Onagraceae genera described in \citet{wagner2007revised} were recovered as monophyletic clades with posterior probabilities of $> 0.95$
+except for sister genera \textit{Neoholmgrenia} and \textit{Camissoniopsis} (posterior = 0.31) (Figure \ref{posteriors}).
+Onagraceae was found to diverge from Lythraceae at 109 Mya (Figure \ref{genera}). 
+Divergence time estimates of other major clades and 95\% highest posterior density (HPD) intervals can be seen in Table \ref{times}.
+
+\begin{table*}
+   \begin{adjustbox}{max width=\textwidth}
+      \begin{tabular}{lllllll}
+         \hline
+         Clade & Mean Age (Mya) & 95\% HPD Min & 95\% HPD Max \\ 
 	 \hline
@@
        Estimated posterior probabilities close to 1.0 are shown in green. All genera described in
-       \citet{wagner2007revised} are monophyletic with posterior probabilities of $> 0.95$
+       \citet{wagner2007revised} were found to be monophyletic with posterior probabilities of $> 0.95$
        except for sister genera \textit{Neoholmgrenia} and \textit{Camissoniopsis} (posterior = 0.31).

### 2014-05-02 manuscript.tex
set up make rtf for word docs
added correlation results

@@
 \usepackage{adjustbox}
-
+\usepackage{natbib}
 \modulolinenumbers[5]
@@
 \begin{table*}
+   \center
    \begin{adjustbox}{max width=\textwidth}
@@
 blah blah
+
+\begin{table*}
+   \center
+   \begin{adjustbox}{max width=\textwidth}
+      \begin{tabular}{l|llllll}
+         \hline
+	 & \multicolumn{4}{c}{Number of Petals} \\
+         Color & 2 & 4 & 5 & 6 \\ 
+	 \hline
+         pink & -0.005 & .011 & ns & ns \\
+	 yellow & -0.008 & .021 & ns & ns \\
+	 white & 0.008 & 0.013 & -0.006 & -0.006 \\
+	 green & ns & -0.011 & ns & ns \\
+	 red & ns & ns & ns & ns \\
+         \hline
+      \end{tabular}
+   \end{adjustbox}
+   \caption{Test statistics for the correlation between flower color and number of petals. 
+	    $d$ values are shown
+            for the pairwise comparison of states with $p< 0.05$. ns indicates no significant
+	    association. The overall $D$ value was 0.263 ($p=0.00$).
+	    }
+   \label{correlations}
+\end{table*}
 

### 2014-05-03 manuscript.tex
fixed up tables, some intro

--- 2014-04-28 manuscript.tex
@@
 \usepackage{adjustbox}
-
+\usepackage{natbib}
+\usepackage{booktabs}
 \modulolinenumbers[5]
@@
 
-blah blah
+Questions: are the taxonomic groups described in \citet{wagner2007revised} monophyletic?
+When did the major clades diverge?
+What are the patterns of petal color and petal number evolution, and
+are their evolution correlated?
+What is the pattern of chromosome number evolution?
 
@@
 \begin{table*}
+   \center
    \begin{adjustbox}{max width=\textwidth}
       \begin{tabular}{lllllll}
-         \hline
+         \toprule
          Group & Age (Mya) & Prior Distribution & Mean & SD & Offset & Reference \\ 
-	 \hline
+	 \midrule
          \textit{Circaea} (Onagraceae) & 12 & lognormal & 0.0 & 2.0 & 12 & \citep{grimsson} \\
@@
          Lythraceae & 82 & lognormal & 0.0 & 2.0 & 82 & \citep{graham} \\
-         \hline
+         \bottomrule
       \end{tabular}
@@
 \begin{table*}
+   \center
    \begin{adjustbox}{max width=\textwidth}
       \begin{tabular}{lllllll}
-         \hline
+         \toprule
          DNA Region & \# of Taxa & Aligned Length & Missing data (\%) & Taxon Coverage Density \\ 
-	 \hline
+	 \midrule
          ITS & 453 & 1746 & 13.2 & 0.87 \\
@@
          R5 & 18 & 3129 & 96.6 & 0.03 \\
-         \hline
+         \bottomrule
       \end{tabular}
@@
 \begin{table*}
+   \center
    \begin{adjustbox}{max width=\textwidth}
       \begin{tabular}{lllllll}
-         \hline
+         \toprule
          Clade & Mean Age (Mya) & 95\% HPD Min & 95\% HPD Max \\ 
-	 \hline
+	 \midrule
          Onagraceae / Lythraceae & 109 & 88 & 131 \\
@@
 	 \textit{Chylismia} / \textit{Oenothera} & 25 & 18 & 31 \\
-         \hline
+         \bottomrule
       \end{tabular}
@@
 blah blah
+
+\begin{table*}
+   \center
+   \begin{adjustbox}{max width=\textwidth}
+      \setlength{\tabcolsep}{20pt}
+      \begin{tabular}{lllllll}
+         \toprule
+	 & \multicolumn{4}{c}{Number of Petals} \\
+         & 2 & 4 & 5 & 6 \\ 
+	 \cmidrule{2-5}
+	 Petal Color \\
+         \hspace{3 mm} Pink & -0.005 & .011 & $ns$ & $ns$ \\
+	 \hspace{3 mm} Yellow & -0.008 & .021 & $ns$ & $ns$ \\
+	 \hspace{3 mm} White & 0.008 & 0.013 & -0.006 & -0.006 \\
+	 \hspace{3 mm} Green & $ns$ & -0.011 & $ns$ & $ns$ \\
+	 \hspace{3 mm} Red & $ns$ & $ns$ & $ns$ & $ns$ \\
+         \bottomrule
+      \end{tabular}
+   \end{adjustbox}
+   \caption{Test statistics for the correlation between flower color and number of petals. 
+	    $d$ values are shown
+            for the pairwise comparison of states with $p< 0.01$. $ns$ indicates no significant
+	    association. The overall $D$ value was 0.263 ($p=0.00$).
+	    }
+   \label{correlations}
+\end{table*}
 

### 2014-05-06 manuscript.tex
revised methods

@@
 
-\title{Fossil-calibrated phylogeny of Onagraceae: patterns of floral and chromosome evolution}
+\title{Fossil-calibrated supermatrix phylogeny of Onagraceae: patterns of floral and chromosome evolution}
 
@@
 
-I downloaded all available DNA sequences from GenBank release 200 PLN division and
-performed an exhaustive all-by-all BLASTn \citep{blast} comparison for sequences in Onagraceae
+Lythraceae was selected to use as an outgroup since previous molecular phylogenetic analyses
+place it sister to Onagraceae \citep{conti1997, Sytsma2004}.
+I downloaded all DNA sequences from GenBank release 200 PLN division and
+performed an exhaustive all-by-all BLASTn \citep{blast} comparison of sequences in Onagraceae
 and Lythraceae.
@@
 two clusters was removed from the supermatrix.
-Code written to assemble the supermatrix is available as the Python module 
-SUMAC \citep[\url{https://github.com/wf8/sumac}]{sumac}, and can be used to assemble
-supermatrices for other taxonomic groups recognized in GenBank.
+The program written to data-mine GenBank and assemble the supermatrix 
+is available as the open source Python module 
+SUMAC \citep{sumac}.
+SUMAC can be used to assemble
+supermatrices for any taxonomic group recognized in GenBank, 
+and is optimized to run on multicore processors and clusters by utilizing multiple parallel processes.
 

### 2014-05-06 manuscript.tex
title change
added chromosome figure and abstract

@@
 
-\usepackage{lineno}
 \usepackage[hidelinks]{hyperref}
@@
 \usepackage{booktabs}
-\modulolinenumbers[5]
+\usepackage{lineno}
 
@@
 
-\title{Fossil-calibrated supermatrix phylogeny of Onagraceae: patterns of floral and chromosome evolution}
+\title{Patterns of floral and chromosome evolution in Onagraceae: a fossil-calibrated supermatrix analysis}
 
@@
 \begin{abstract}
-blah blah
+Onagraceae is a cosmopolitan plant family in the order Myrtales comprised of 22 genera and 657 species.
+I used a GenBank data-mining approach to construct an 11 gene supermatrix of 521 taxa in Onagraceae and Lythraceae.
+Maximum likelihood and Bayesian inference phylogenetic analyses were performed, 
+and divergence time estimates were calibrated using 5 fossils.
+Chromosome number evolution was estimated using probabilistic models,
+and ancestral states of petal color and petal number were reconstructed using maximum likelihood.
+Correlated evolution between petal color and petal number was tested using
+stochastic character mapping.
+The monophyly of all major clades in Onagraceae was supported,
+and Onagraceae was estimated to have diverged from Lythraceae 109 Mya. 
+The base chromosome number of Onagraceae was inferred to be $x=5$.
+Petal color and petal number were found
+to follow a pattern of correlated evolution ($p=0.00$), 
+suggesting concerted shifts in floral traits may play an important role shaping diversification of Onagraceae. 
 \end{abstract}
@@
 
-\linenumbers
+%\modulolinenumbers[5]
+%\linenumbers
 
@@
 with the Markov k-state 1 parameter (Mk1) model \citep{lewis2001likelihood}.
+Ancestral chromosome numbers were inferred using maximum likelihood and Bayesian methods 
+as implemented in ChromEvol 2.0 \citep{chromevol}.
+Eight different models of chromosome evolution were fit to the Bayesian MCC phylogeny using ChromEvol,
+and the best fit model was selected using Akaike's information criterion \citep{aic}.
 
@@
 
+\begin{figure*}[p]
+    \vspace*{-2cm}
+    \makebox[\linewidth]{
+        \includegraphics[width=1.6\linewidth, trim=0 0 0 0, clip=true]{chromtree_labeled}
+    }
+    \caption{
+    Maximum likelihood estimates of ancestral chromosome numbers over the Bayesian MCC phylogeny
+    using the 4 parameter ($\lambda$, $\delta$, $\mu$, $\rho$) model of chromosome evolution selected by Akaike's information criterion.
+    The base number of Onagraceae is inferred to be $x=5$.
+    }
+    \label{chromtree}
+\end{figure*}
+
 %%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-06 manuscript.tex
stochastic character mapping to results

@@
 Correlated evolution between petal color and petal number was tested using
-stochastic character mapping.
+Bayesian stochastic character mapping.
 The monophyly of all major clades in Onagraceae was supported,
@@
 Character data was assembled from the comprehensive \citet{wagner2007revised} Onagraceae monograph.
+Ancestral chromosome numbers were inferred using maximum likelihood and Bayesian methods 
+as implemented in ChromEvol 2.0 \citep{chromevol}.
+Eight different models of chromosome evolution were fit to the Bayesian MCC phylogeny using ChromEvol,
+and the best fit model was selected using Akaike's information criterion \citep{aic}.
 Ancestral character state reconstructions of petal number and petal color were performed using Mesquite v2.75 \citep{mesquite}
@@
 with the Markov k-state 1 parameter (Mk1) model \citep{lewis2001likelihood}.
-Ancestral chromosome numbers were inferred using maximum likelihood and Bayesian methods 
-as implemented in ChromEvol 2.0 \citep{chromevol}.
-Eight different models of chromosome evolution were fit to the Bayesian MCC phylogeny using ChromEvol,
-and the best fit model was selected using Akaike's information criterion \citep{aic}.
+Additionally, Bayesian stochastic character mapping \citep{Huelsenbeck2003} was used to test whether
+petal color and petal number covaried over the phylogeny. 
+Models of petal color and petal number evolution
+were configured in the program SIMMAP v1.5 \citep{bollback2006simmap}
+using unordered states, gamma rate priors, and equal state bias priors.
+The correlation analysis used SIMMAP's default
+predictive sampling configuration
+and calculated the following correlation statistics: 
+$D$ the overall association between
+the two characters, and $d_{ij}$ the association between the
+individual states of each character \citep{Huelsenbeck2003}.
 
@@
    \end{adjustbox}
-   \caption{Test statistics for the correlation between flower color and number of petals. 
+   \caption{Test statistics for the correlation between flower color evolution and petal number evolution. 
 	    $d$ values are shown

### 2014-05-07 manuscript.tex
added much to results

@@
 
-\paragraph{Character evolution}
-blah blah
+\paragraph{Chromosome number evolution}
+Out of the eight models of chromosome evolution tested by ChromEvol,
+the four parameter $CONST\_RATE\_DEMI\_EST$ model was the best fit according to the AIC scores. 
+This model has separate rate parameters for
+chromosome gains ($\lambda$),
+losses ($\delta$),
+demi-polyploid events ($\mu$),
+and polyploid events ($\rho$).
+Demi-polyploid events occur when reduced and unreduced gametes unite
+resulting in a $3x$ cytotype.
+The base chromosome number of Onagraceae was estimated to be $x=5$,
+and the ancestral number of the large genus \textit{Epilobium} was estimated to be $n=9$,
+though the extant taxa are split in two large monophyletic clades of 
+$n=10$ and $n=18$, respectively (Figure \ref{chromtree}).
+
+\paragraph{Floral evolution}
+The maximum likelihood estimate of petal color for the 
+last common ancestor of Onagraceae was pink 
+with a proportional likelihood ($PL$) of $0.61$.
+White was the second most likely estimate ($PL=0.11$).
+The maximum likelihood estimate of petal number for the
+last common ancestor of Onagraceae was 4 ($PL=1.0$), with 15
+independent transitions to 0, 2, 5, or 6 petals (Figure XXXX).
+Petal color and petal number were found to significantly covary over the 
+phylogeny ($D=0.263, p=0.00$). Multiple signficant correlations
+($d_{ij}$) were found between individual states of each character
+and are summarized in Table \ref{correlations}.
 
@@
    \end{adjustbox}
-   \caption{Test statistics for the correlation between flower color evolution and petal number evolution. 
-	    $d$ values are shown
+   \caption{Test statistics for the correlation between petal color evolution and petal number evolution. 
+	    $d_{ij}$ values are shown
             for the pairwise comparison of states with $p< 0.01$. $ns$ indicates no significant

### 2014-05-07 manuscript.tex
minor copy edits

@@
 and divergence time estimates were calibrated using 5 fossils.
-Chromosome number evolution was estimated using probabilistic models,
+Chromosome number evolution was inferred using probabilistic models,
 and ancestral states of petal color and petal number were reconstructed using maximum likelihood.
@@
 to follow a pattern of correlated evolution ($p=0.00$), 
-suggesting concerted shifts in floral traits may play an important role shaping diversification of Onagraceae. 
+suggesting concerted shifts in floral traits may play an important role in shaping diversification of Onagraceae. 
 \end{abstract}
@@
 
-Lythraceae was selected to use as an outgroup since previous molecular phylogenetic analyses
+Lythraceae was selected as an outgroup since previous molecular phylogenetic analyses
 place it sister to Onagraceae \citep{conti1997, Sytsma2004}.
@@
 SUMAC \citep{sumac}.
-SUMAC can be used to assemble
+SUMAC will assemble
 supermatrices for any taxonomic group recognized in GenBank, 
@@
 so only the Bayesian MCC tree (Figures \ref{posteriors} and \ref{genera}) is shown here.
-All Onagraceae genera described in \citet{wagner2007revised} were recovered as monophyletic clades with posterior probabilities of $> 0.95$
+All Onagraceae genera described in \citet{wagner2007revised} were recovered as monophyletic clades with posterior probabilities $> 0.95$
 except for sister genera \textit{Neoholmgrenia} and \textit{Camissoniopsis} (posterior = 0.31) (Figure \ref{posteriors}).
@@
             for the pairwise comparison of states with $p< 0.01$. $ns$ indicates no significant
-	    association. The overall $D$ value was 0.263 ($p=0.00$).
+	    association. Negative $d_{ij}$ values express negative correlation.
+	    The overall $D$ value was 0.263 ($p=0.00$).
 	    }

### 2014-05-07 manuscript.tex
added mesquite figures

@@
 last common ancestor of Onagraceae was pink 
-with a proportional likelihood ($PL$) of $0.61$.
+with a proportional likelihood ($PL$) of $0.61$ (Figure \ref{petal_color}).
 White was the second most likely estimate ($PL=0.11$).
@@
 last common ancestor of Onagraceae was 4 ($PL=1.0$), with 15
-independent transitions to 0, 2, 5, or 6 petals (Figure XXXX).
+independent transitions to 0, 2, 5, or 6 petals (Figure \ref{petal_number}).
 Petal color and petal number were found to significantly covary over the 
@@
 
+\begin{figure*}[p]
+    \vspace*{-2cm}
+    \makebox[\linewidth]{
+        \includegraphics[width=1.6\linewidth, trim=0 0 0 0, clip=true]{petal_color}
+    }
+    \caption{
+      Maximum likelihood ancestral state reconstruction of petal color.
+      Black represents absence of petals. Vertical time scale is in millions of years before present.
+    }
+    \label{petal_color}
+\end{figure*}
+
+\begin{figure*}[p]
+    \vspace*{-2cm}
+    \makebox[\linewidth]{
+        \includegraphics[width=1.6\linewidth, trim=0 0 0 0, clip=true]{petal_number}
+    }
+    \caption{
+    Maximum likelihood ancestral state reconstructions of petal number. 
+    White = 0, Blue = 2, Green = 4, Yellow = 5, Red = 6. Vertical time scale is in millions of years before present.
+    }
+    \label{petal_number}
+\end{figure*}
+
 %%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-08 manuscript.tex
fixed figure

@@
          rps16 & 74 & 1016 & 85.8 & 0.14 \\
-         rbcL & 64 & 1310 & 87.7 & 0.12 \\
+         rbcL2 & 64 & 1310 & 87.7 & 0.12 \\
          PgiC2 & 47 & 4028 & 91.0 & 0.09 \\

### 2014-05-10 manuscript.tex
fixed errors in mesquite character coding

@@
 last common ancestor of Onagraceae was pink 
-with a proportional likelihood ($PL$) of $0.61$ (Figure \ref{petal_color}).
-White was the second most likely estimate ($PL=0.11$).
+with a proportional likelihood ($PL$) of $0.56$ (Figure \ref{petal_color}).
+White was the second most likely estimate ($PL=0.19$).
 The maximum likelihood estimate of petal number for the
-last common ancestor of Onagraceae was 4 ($PL=1.0$), with 15
-independent transitions to 0, 2, 5, or 6 petals (Figure \ref{petal_number}).
+last common ancestor of Onagraceae was 4 ($PL=1.0$), with 8
+independent transitions to 0, 2, or 5 petals (Figure \ref{petal_number}).
 Petal color and petal number were found to significantly covary over the 
@@
     Maximum likelihood ancestral state reconstructions of petal number. 
-    White = 0, Blue = 2, Green = 4, Yellow = 5, Red = 6. Vertical time scale is in millions of years before present.
+    White = 0, Blue = 2, Green = 4, Orange = 5. Vertical time scale is in millions of years before present.
     }

### 2014-05-14 manuscript.tex
lots of intro work

@@
 suggesting concerted shifts in floral traits may play an important role in shaping diversification of Onagraceae. 
+Further analyses should be performed to test whether shifts in
+floral traits are correlated with pollinator shifts,
+and to determine how these shifts affect diversification rates.
+Furthermore, this work should be extended to test historical biogeographical hypotheses
+that explain the intercontinental disjunct distribution of Onagraceae.
 \end{abstract}
@@
 
-Questions: are the taxonomic groups described in \citet{wagner2007revised} monophyletic?
-When did the major clades diverge?
-What are the patterns of petal color and petal number evolution, and
-are their evolution correlated?
-What is the pattern of chromosome number evolution?
+Introduce supermatrix analyses....
+
+Onagraceae, the evening primroses, are a well-studied family
+of flowering plants with 22 described genera and around 660 species \citep{wagner2007revised}.
+The family is found on all continents except Antartica,
+and pollen fossils date Onagraceae to the Upper Cretaceous \citep{grimsson}.
+Onagraceae is well placed in the order Myrtales,
+and previous molecular phylogenetic studies find it most closely related to Lythraceae \citep{conti1997, Sytsma2004}.
+Though there have been many genus-level molecular phylogenetic studies in Onagraceae
+\citep{Berry2004, Evans2009, Hoggard2004, Xie2009, Baum1994, Wagner2005}, 
+the most recent family-level analyses \citep{Levin2003, Levin2004}
+used only one or two exemplars from some genera. 
+Furthermore, the historical biogeography, divergence times of the major clades, and 
+family-level patterns and processes of diversification remain poorly understood.
+With the wealth of data available, Onagraceae is an
+ideal group to apply a data-mining supermatrix approach
+with the goal of constructing a densely sampled family-level phylogeny.
+
+Here, I use a well-resolved supermatrix phylogeny of Onagraceaea to
+answer qestions such as:
+1) Are the taxonomic groups described in \citet{wagner2007revised} monophyletic?
+2) When did the major clades diverge?
+3) What is the pattern of chromosome number
+evolution?
+4) What are the patterns of petal color and petal number evolution, and are their 
+evolution correlated?
+Finally, I also discuss ways in which these analyses can be extended
+to further examine the patterns and processes of
+evolution in Onagraceae.
 
@@
 \paragraph{Phylogeny and divergence time estimates}
-The topologies of the ML and Bayesian phylogenies were identical for all major clades within Onagraceae, 
+The topologies of the ML and Bayesian phylogenies were nearly identical for all major clades within Onagraceae, 
 so only the Bayesian MCC tree (Figures \ref{posteriors} and \ref{genera}) is shown here.
-All Onagraceae genera described in \citet{wagner2007revised} were recovered as monophyletic clades with posterior probabilities $> 0.95$
+The major exception is that in the ML tree the genus \textit{Chamerion} was recovered
+as sister to \textit{Epilobium}, whereas in the MCC tree XXXXXX.......
+In the MCC tree all Onagraceae genera described in \citet{wagner2007revised} were recovered as monophyletic clades with posterior probabilities $> 0.95$
 except for sister genera \textit{Neoholmgrenia} and \textit{Camissoniopsis} (posterior = 0.31) (Figure \ref{posteriors}).
@@
       \setlength{\tabcolsep}{20pt}
-      \begin{tabular}{lllllll}
+      \begin{tabular}{llllll}
          \toprule
-	 & \multicolumn{4}{c}{Number of Petals} \\
-         & 2 & 4 & 5 & 6 \\ 
-	 \cmidrule{2-5}
+	 & \multicolumn{3}{c}{Number of Petals} \\
+         & 2 & 4 & 5 \\ 
+	 \cmidrule{2-4}
 	 Petal Color \\
-         \hspace{3 mm} Pink & -0.005 & .011 & $ns$ & $ns$ \\
-	 \hspace{3 mm} Yellow & -0.008 & .021 & $ns$ & $ns$ \\
-	 \hspace{3 mm} White & 0.008 & 0.013 & -0.006 & -0.006 \\
-	 \hspace{3 mm} Green & $ns$ & -0.011 & $ns$ & $ns$ \\
-	 \hspace{3 mm} Red & $ns$ & $ns$ & $ns$ & $ns$ \\
+         \hspace{3 mm} Pink & -0.005 & .011 & $ns$ \\
+	 \hspace{3 mm} Yellow & -0.008 & .021 & $ns$ \\
+	 \hspace{3 mm} White & 0.008 & 0.013 & -0.006 \\
+	 \hspace{3 mm} Green & $ns$ & -0.011 & $ns$ \\
+	 \hspace{3 mm} Red & $ns$ & $ns$ & $ns$ \\
          \bottomrule
@@
 %%%%%%%%%%%%%%%%%%%%%%%
-%% conclusion
+%% discussion
 %%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Conclusion}
+\section{Discussion}
 
-blah blah
+discuss monophyly of tribe Epilobieae, which was supported in 
+maximum likelihood analysis but not supported in Bayesian MCC, could be due to
+fossil constraints.....
 

### 2014-05-14 manuscript.tex
removed elsevier preprint
intro work

@@
 \documentclass[review]{elsarticle}
+
+% to remove elsearticle 'preprint journal'
+\makeatletter
+\def\ps@pprintTitle{%
+ \let\@oddhead\@empty
+ \let\@evenhead\@empty
+ \def\@oddfoot{\centerline{\thepage}}%
+ \let\@evenfoot\@oddfoot}
+\makeatother
+%%\journal{IB200}
 
@@
 \graphicspath{ {./figures/} }
-
-\journal{IB200}
 
@@
 
-Introduce supermatrix analyses....
+Biologists are increasingly using massive phylogenetic
+datasets to reconstruct ever-growing portions of the Tree of Life 
+in pursuit of large-scale evolutionary questions.
+A common approach to building large phylogetic trees has been
+the supermatrix approach, in which multiple datasets are
+combined and analyzed simultaneously \citep{DeQueiroz2007}.
+However, assembling and utilizing supermatrices is not trivial
+due to difficulties such as determining homology of molecular
+sequences, defining operational taxonomic units, 
+and managing the amount of missing data \citep{Roure2007, Yan2005}.
+phlawd and phylota...
+Supermatrix approaches have been successfully used in a number
+of large scale angiosperm evolutionary studies including, bling, blang, blot.
+In this study I use a supermatrix to build a phylogeny of
+the family Onagraceae, and implement a novel method for supermatrix
+construction in a Python package.
 
-Onagraceae, the evening primroses, are a well-studied family
+Onagraceae, the evening primrose family, is a well-studied family
 of flowering plants with 22 described genera and around 660 species \citep{wagner2007revised}.
@@
 and pollen fossils date Onagraceae to the Upper Cretaceous \citep{grimsson}.
-Onagraceae is well placed in the order Myrtales,
+Onagraceae falls within the order Myrtales,
 and previous molecular phylogenetic studies find it most closely related to Lythraceae \citep{conti1997, Sytsma2004}.
@@
 family-level patterns and processes of diversification remain poorly understood.
-With the wealth of data available, Onagraceae is an
+With over 5000 DNA sequences available in GenBank from previous studies, Onagraceae is an
 ideal group to apply a data-mining supermatrix approach
@@
 
-Here, I use a well-resolved supermatrix phylogeny of Onagraceaea to
-answer qestions such as:
-1) Are the taxonomic groups described in \citet{wagner2007revised} monophyletic?
-2) When did the major clades diverge?
-3) What is the pattern of chromosome number
-evolution?
-4) What are the patterns of petal color and petal number evolution, and are their 
+Here, I use a well-resolved supermatrix phylogeny of Onagraceae to
+test whether the taxonomic groups described in \citet{wagner2007revised} are monophyletic,
+and to answer the following questions:
+1) When did the major clades diverge?
+2) What is the pattern of chromosome number evolution?
+3) What are the patterns of petal color and petal number evolution, and are their 
 evolution correlated?
-Finally, I also discuss ways in which these analyses can be extended
-to further examine the patterns and processes of
-evolution in Onagraceae.
+Lastly, I discuss ways in which these analyses can be extended
+to further examine the evolution of Onagraceae, 
+and describe some recently developed metrics that can
+be used to assess the informativeness of phylogenetic supermatrices.
 

### 2014-05-15 manuscript.tex
rough draft intro complete

--- 2014-05-14 manuscript.tex
@@
 
-Introduce supermatrix analyses....
+In pursuit of large-scale evolutionary questions,
+biologists are increasingly using massive phylogenetic
+datasets to reconstruct ever-growing portions of the tree of life. 
+A common approach to building large phylogenetic trees has been
+the supermatrix approach, in which multiple datasets are
+combined and analyzed simultaneously \citep{DeQueiroz2007}.
+However, assembling and utilizing supermatrices is not trivial
+due to challenges such as determining homology of molecular
+sequences, assembling chimeric operational taxonomic units, 
+limiting the amount of missing data,
+and managing data integrity and computational feasibility \citep{Roure2007, Yan2005}.
+Despite these difficulties, considerable bioinformatic advances 
+have made supermatrix construction more common,
+particularly in large-scale evolutionary plant studies.
+For example, \citet{Smith2011}
+inferred a phylogeny of 55473 seed plants,
+\citet{Hinchliff2013} estimated the phylogeny of 1500 members of Cyperaceae,
+and \citet{Sarkinen2013} built a time calibrated tree of Solanaceae with 1076 tips.
 
-Onagraceae, the evening primroses, are a well-studied family
+With over 5000 DNA sequences available in GenBank from previous studies, 
+the plant family Onagraceae is an
+ideal group to apply a data-mining supermatrix approach.
+Onagraceae, the evening primrose family, is a well-studied 
+monophyletic clade
 of flowering plants with 22 described genera and around 660 species \citep{wagner2007revised}.
@@
 Furthermore, the historical biogeography, divergence times of the major clades, and 
-family-level patterns and processes of diversification remain poorly understood.
-With the wealth of data available, Onagraceae is an
-ideal group to apply a data-mining supermatrix approach
-with the goal of constructing a densely sampled family-level phylogeny.
+family-level patterns and processes of diversification remain poorly understood
+and untested over a molecular phylogeny.
 
-Here, I use a well-resolved supermatrix phylogeny of Onagraceaea to
-answer qestions such as:
-1) Are the taxonomic groups described in \citet{wagner2007revised} monophyletic?
-2) When did the major clades diverge?
-3) What is the pattern of chromosome number
-evolution?
-4) What are the patterns of petal color and petal number evolution, and are their 
+Here, I use a well-resolved and densely-sampled supermatrix phylogeny of Onagraceae to
+test whether the taxonomic groups described in \citet{wagner2007revised} are monophyletic,
+and to answer the following questions:
+1) When did the major clades diverge?
+2) What is the pattern of chromosome number evolution?
+3) What are the patterns of petal color and petal number evolution, and are their 
 evolution correlated?
-Finally, I also discuss ways in which these analyses can be extended
-to further examine the patterns and processes of
-evolution in Onagraceae.
+Lastly, I consider ways in which these analyses can be extended
+to further examine the evolution of Onagraceae, 
+and discuss the novel method used here to construct the supermatrix.
 
@@
 maximum likelihood analysis but not supported in Bayesian MCC, could be due to
-fossil constraints.....
+fossil constraints on genus Epilobium, it may be that the fossil described should
+have constrained Epilobium + Chamerion due to taxonomic changes.....
+need to check fossil synapomorphies to see where it should be placed!
+
+Petal color and petal number were found to follow a pattern 
+of correlated evolution (p = 0.00), 
+suggesting concerted shifts in floral traits 
+may play an important role in shaping diversification of Onagraceae. 
+Further analyses should be performed to test 
+whether shifts in floral traits are correlated with pollinator shifts, 
+and to determine how these shifts affect diversification rates. 
+Furthermore, this work should be extended to test historical 
+biogeographical hypotheses that explain the intercontinental disjunct distribution of Onagraceae.
+
+For example, the PhyLoTA Browser 
+provides a web interface to view all GenBank sequences within
+taxonomic groups clustered into homologs \citep{Sanderson2008}.
+A different approach is implemented in the program PHLAWD,
+which constructs clusters of homologs around
+guide sequences provided by the user \citep{Smith2009}.
+However, the method used by SUMAC combines both approaches;
+the user can perform an exploratory clustering of all GenBank sequences
+within a taxonomic group, and also provide guide sequences to build
+homologous clusters in a more targeted manner. SUMAC provides
+a framework with which GenBank can be repeatedly mined 
+using different taxonomic groups and/or different guide sequences
+and the resulting data matrices can be compared.
+
+describe some recently developed metrics that can
+be used to assess the informativeness of phylogenetic supermatrices.
 

### 2014-05-16 manuscript.tex
revisions and discussion

@@
 Bayesian stochastic character mapping.
-The monophyly of all major clades in Onagraceae was supported,
+The monophyly of all genera in Onagraceae was supported,
 and Onagraceae was estimated to have diverged from Lythraceae 109 Mya. 
@@
 combined and analyzed simultaneously \citep{DeQueiroz2007}.
-However, assembling and utilizing supermatrices is not trivial
-due to challenges such as determining homology of molecular
+However, assembling and utilizing supermatrices is challenging
+due to difficulties such as determining homology of molecular
 sequences, assembling chimeric operational taxonomic units, 
@@
 and managing data integrity and computational feasibility \citep{Roure2007, Yan2005}.
-Despite these difficulties, considerable bioinformatic advances 
+Despite these challenges, considerable bioinformatic advances 
 have made supermatrix construction more common,
@@
 
-With over 5000 DNA sequences available in GenBank from previous studies, 
-the plant family Onagraceae is an
-ideal group to apply a data-mining supermatrix approach.
+The plant family Onagraceae is an
+ideal group to apply a data-mining supermatrix approach
+due to the over 5000 DNA sequences available in GenBank from previous studies.
 Onagraceae, the evening primrose family, is a well-studied 
@@
 Here, I use a well-resolved and densely-sampled supermatrix phylogeny of Onagraceae to
-test whether the taxonomic groups described in \citet{wagner2007revised} are monophyletic,
-and to answer the following questions:
-1) When did the major clades diverge?
-2) What is the pattern of chromosome number evolution?
-3) What are the patterns of petal color and petal number evolution, and are their 
-evolution correlated?
+test the monophyly of the taxonomic groups described in \citet{wagner2007revised}
+and estimate when they diverged.
+I then inferred the pattern of chromosome number evolution using probabilistic models and
+examined the patterns of petal color and petal number evolution, 
+including a correlation test of the evolution of these two floral traits.
 Lastly, I consider ways in which these analyses can be extended
 to further examine the evolution of Onagraceae, 
-and discuss the novel method used here to construct the supermatrix.
+and discuss the novel method used here to construct the phylogenetic supermatrix.
 
@@
 so only the Bayesian MCC tree (Figures \ref{posteriors} and \ref{genera}) is shown here.
-The major exception is that in the ML tree the genus \textit{Chamerion} was recovered
-as sister to \textit{Epilobium}, whereas in the MCC tree XXXXXX.......
-In the MCC tree all Onagraceae genera described in \citet{wagner2007revised} were recovered as monophyletic clades with posterior probabilities $> 0.95$
+The only exception is the tribe Epilobieae \textit{sensu} \citet{wagner2007revised}; 
+the ML analysis recovered tribe Epilobieae as monophyletic, 
+whereas according to the Bayesian estimate it is paraphyletic.
+All Onagraceae genera described in \citet{wagner2007revised} were recovered as monophyletic clades 
+in the Bayesian MCC with posterior probabilities $> 0.95$,
 except for sister genera \textit{Neoholmgrenia} and \textit{Camissoniopsis} (posterior = 0.31) (Figure \ref{posteriors}).
@@
 Petal color and petal number were found to significantly covary over the 
-phylogeny ($D=0.263, p=0.00$). Multiple signficant correlations
+phylogeny ($D=0.263, p=0.00$). Multiple significant correlations
 ($d_{ij}$) were found between individual states of each character
@@
 
-discuss monophyly of tribe Epilobieae, which was supported in 
-maximum likelihood analysis but not supported in Bayesian MCC, could be due to
-fossil constraints on genus Epilobium, it may be that the fossil described should
-have constrained Epilobium + Chamerion due to taxonomic changes.....
-need to check fossil synapomorphies to see where it should be placed!
+The paraphyly of tribe Epilobieae in the Bayesian MCC tree (Figure \ref{genera})
+conflicts with both the 
+maximum likelihood analysis (not shown) and the previous studies of \citet{Levin2003, Levin2004},
+all of which found tribe Epilobieae to be monophyletic.
+This incongruence is likely due to the placement of the
+fossil constraint on the genus \textit{Epilobium} used in the Bayesian analysis. 
+In this study I followed the taxonomy of \citet{wagner2007revised}, 
+which splits tribe Epilobieae into the sister genera \textit{Epilobium}
+and \textit{Chamerion}.
+However, the pollen morphology descriptions used by \citet{grimsson} to identify the
+\textit{Epilobium} fossil did not recognize \textit{Chamerion} as separate from
+\textit{Epilobium} \citep{Punt2003, Patel1984}. 
+The divergence times will be reestimated using the \textit{Epilobium} fossil 
+to constrain both genera. 
+This will likely result in topological changes to the Bayesian MCC, 
+which means that the 
+trait evolution estimates presented here should be considered preliminary.
 
-Petal color and petal number were found to follow a pattern 
+The reconstruction of ancestral chromosome numbers 
+validates many estimates made by \citet{Raven1979}.
+Raven's estimates of the base chromosome number for the genera
+\textit{Ludwigia}, \textit{Fuchsia}, \textit{Circaea}, \textit{Hauya}, 
+\textit{Epilobium}, \textit{Camissonia}, and \textit{Oenothera}
+all agree with the reconstruction presented here (Figure \ref{chromtree}).
+However, for two genera I report estimates that differ from Raven's: 
+for \textit{Clarkia} $x=9$ and for \textit{Lopezia} $x=10$.
+Furthermore, \citet{Raven1979} proposed that the ancestral
+chromosome number of all Onagraceae was $x=11$, whereas
+my estimate suggests $x=5$. Further testing with the ChromEvol
+models should be carried out, particularly to see if the inclusion
+of the outgroup (Lythraceae) affected the estimate of the
+base chromosome number of Onagraceae.
+
+The analysis of petal color and petal number evolution using
+Bayesian stochastic character mapping
+revealed that the two floral traits followed a pattern 
 of correlated evolution (p = 0.00), 
@@
 may play an important role in shaping diversification of Onagraceae. 
-Further analyses should be performed to test 
-whether shifts in floral traits are correlated with pollinator shifts, 
-and to determine how these shifts affect diversification rates. 
-Furthermore, this work should be extended to test historical 
+To test this, further analyses will be performed to 
+determine if these shifts in floral morphology affect diversification rates.
+Other floral traits, such as corolla width and diurnal or nocturnal flower opening, 
+could be included in future analyses.
+Tests will also be performed to evaluate whether 
+shifts in floral traits are correlated with pollinator shifts.
+These analyses could provide evidence that pollinators are driving the
+evolution of floral morphology in Onagraceae \citep{DeWittSmith2010}.
+
+This work will also be extended to test historical 
 biogeographical hypotheses that explain the intercontinental disjunct distribution of Onagraceae.
+The family-level supermatrix phylogeny will be used to infer the location of the 
+last common ancestor of Onagraceae, which has been hypothesized
+to have been in South America \citep{raven1974angiosperm, raven1988onagraceae}.
+The biogeographical analysis will also test the suggestion made by \citet{Berry2004} that \textit{Fuchsia}
+originated in Central America. 
+To my knowledge this will be the first reconstruction of 
+the historical biogeography of Onagraceae over a molecular phylogeny.
 
+The strength of the supermatrix approach to phylogenetic estimation
+is that it combines data from diverse sources into one large analysis.
+This strength is amplified by data-mining GenBank in an exploratory
+fashion, so that sequence data not necessarily used in previous systematic
+studies can also be incorporated into the supermatrix.
+A number of software tools for building supermatrices are already available 
+that use methods similar to those performed here.
 For example, the PhyLoTA Browser 
@@
 guide sequences provided by the user \citep{Smith2009}.
-However, the method used by SUMAC combines both approaches;
+However, the method used here (implemented in SUMAC) combines elements of both approaches;
 the user can perform an exploratory clustering of all GenBank sequences
 within a taxonomic group, and also provide guide sequences to build
-homologous clusters in a more targeted manner. SUMAC provides
-a framework with which GenBank can be repeatedly mined 
-using different taxonomic groups and/or different guide sequences
+homologous clusters in a more targeted manner. 
+SUMAC is being extended to calculate metrics that assess
+the informativeness of phylogenetic supermatrices 
+that are unavailable in other software tools.
+For example, the next version of SUMAC will calculate
+an index of substitution saturation based on information entropy \citep{Xia2003}
+for each alignment,
+and also calculate the partial phylogenetic decisiveness 
+of the supermatrix,
+which measures the percentage of trees (out of all possible trees)
+that can be inferred given the data missing from the supermatrix \citep{Sanderson2010, Fischer2012}.
+By calculating these metrics SUMAC will provide a unique
+framework with which GenBank can be repeatedly mined 
+using different settings
 and the resulting data matrices can be compared.
-
-describe some recently developed metrics that can
-be used to assess the informativeness of phylogenetic supermatrices.
+With further methodological refinement, 
+supermatrix approaches will be indispensable
+to future large-scale phylogenetic studies.
 
